---
title: "Simple R Functions "
author: "yaru peng"
date: "January 26, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}

## simple example
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b

```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2){
  
  n = length(xVec2)
  
  return(xVec2^(1:n)/(1:n))
}



c <- tmpFn2(a)
c


```

######  
  
  
  
(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$
```{r}
tmpFn3 <- function(x,n){
  1+sum((x^(1:n))/(1:n))
}

d <-tmpFn3(1,10)
d
```
\newpage
### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$

Try out your function.  \texttt{tmpFn}(c(1:5,6:1))
```{r}
tempFn <- function(xVec){
  x<-length(xVec)
  (xVec[1:(x-2)]+xVec[2:(x-1)]+xVec[3:x])/3
}

tempFn(c(1:5,6:1))
```
### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.
```{r}
tmpFn <-function(xVec){
  ifelse(xVec<0,xVec^2+2*xVec+3,ifelse(xVec<2,xVec+3,xVec^2+4*xVec-7))
}

curve(tmpFn, from=-3, to=3)
            
```
### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$
```{r}
newMatrix<-function(x){
  x[x%%2==1]<-2*x[x%%2==1]
}
```

### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$
```{r}

tFn<-function(n,k){
  newMat<-matrix(rep(0,n*n), byrow = TRUE)
  newMat <- diag(k,nrow=n)
  newMat[abs(row(newMat)-col(newMat))==1]<-1
  newMat
}
```

### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.
```{r}
quadrant <-function(alpha){
  floor(alpha/90)%%4+1
}
```

### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1`963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.
```{r}
weekday<-function(day,month,year){
  month <- month-2
  if (month<=0){
    month <- month+12
    year <- year-1
  }
 century<-as.integer(substring(as.character(year*100),1,2))
 year <- year%%100
 vcs <-floor(2.6*month-0.2)+day+year+year%/%4+century%/%4-2*century
 c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")[1+vcs%%7]
}
weekday(03,14,1996)
```

(b) Does your function work if the input parameters \text{day, month, and year} are vectors with the same length and valid entries?
```{r}
weekday2<-function(day,month,year){
  new<-month<=2
  month<-month-2+12*new
  year<-year-new
  century<-as.integer(substring(as.character(year*100),1,2))
 year <- year%%100
 vcs <-floor(2.6*month-0.2)+day+year+year%/%4+century%/%4-2*century
 c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")[1+vcs%%7]
 new
 }
```

### 8.
(a)
```{r}
testloop<-function(n){
  x<-rep(NA,n-1)
    x[1]<-1
    x[2]<-2
  for(j in 3:(n-1))
    x[j]<-x[j-1]+2/x[j-1]
  x
}
```

(b)
```{r}
testloop2<-function(yVec){
  n<-length(yVec)
  if(n<=0){0}
  else{sum(exp(1:n))}
}
```

###9.
(a)
```{r}
quadmap<-function(start,rho,niter){
  x <- rep(NA,niter)
   x[1]<-start
  for(n in 1:(niter-1)){
    x[n+1]<-rho*x[n]*(1-x[n])
  }
  x
  }
tmp <-quadmap(start=0.95,rho=2.99,niter=500)
```
(b)
```{r}
quadmap2<-function(start,rho){
  x1 <- start
  x2 <- rho*x1*(1-x1)
  n <- 1
  while(abs(x2-x1)>=0.02){
    x1 <- x2
    x2 <- rho*x1*(1-x1)
    n<- n+1 
  }
  n
}
tmp<-quadmap2(0.95,2.99)
```

###10
(a)
```{r}
   tmpFn<-function(xVec){
     len <- length(xVec)
     diff<- xVec-mean(xVec)
     r1 <- sum(diff[2:len]*diff[1:(len-1)])/sum(diff^2)
     r2 <- sum(diff[3:len]*diff[1:(len-2)])/sum(diff^2)
     list(r1=r1,r2=r2)
   }
   vecs<-seq(2, 56, by=3)
   tmpFn(vecs)
```
(b)
```{r}
   tmpFn<-function(xVec,k){
     len<-length(xVec)
     diff<-xVec-mean(xVec)
     div<-sum(diff^2)
     tmpfn<-function(i){sum(diff[(i+1):len]*diff[1:(len-i)])/div}
     c(1,sapply(1:k,tmpfn))
   }
   vecs<-seq(2,56,by=3)
   tmpFn(vecs,3)
```



